Union and intersection contracts are hard, actually

### 2021-06-10 paper.tex
Small stylistic changes
Section 5.4

Racket is a language based on the Scheme dialect of Lisp. Among established
languages, Racket is probably the one with the most comprehensive contract
system\cite{RacketContracts}. Regarding union and intersections, Racket provides
the \racket{and/c} and \racket{or/c} combinators for contract.  These
combinators correspond to the naive interpretation of unions and intersections
(as in Figure~\ref{fig:naive-impl} for intersections): \racket{and/c} (resp.  \racket{or/c})
simply checks sequentially that all (resp. one) contracts pass.  In consequence,
unions or intersections of higher-order contracts don't match intuition, and in
particular intersections don't model overloading. As in
Figure~\ref{fig:naive-impl}, the example given in
Figure~\ref{fig:racket-overloading} always fail because no argument satisfies
both \racket{number?} and \racket{string?}.

% Among those contracts, they provide \racket{and/c} and \racket{or/c} contracts.
% These contracts are lightweight versions of intersections and unions, they simply
% check that every contract is valid, in the case of \racket{and/c},
% and that at least one contract
% is valid, in the case of \racket{or/c}.

% So, for instance, the usual usecase of intersection to encode bounded overloading
% does not work on Racket. Take a look at the next example, the function
% \racket{overload} should accept a call like \racket{(overload 3)}, but it fails
% since 3 is not a valid \racket{string?}.

@@

When higher-order contracts are combined using \racket{or/c}, Racket imposes
that contracts must be distinguishable by their arity. Doing so, there is at most
one candidate that can be selected directly. This is illustrated in
Figure~\ref{code:racket:or/c:working} whose program is accepted. On the other
hand, the program of Figure~\ref{code:racket:or/c:non-working} is rejected.

% that could (potentially) apply to a function, these have to be
% differentiated by first order markers.
% For instance, the code on Figure~\ref{code:racket:or/c:working}
% would correctly allow only to call
% \racket{united} with a \racket{number?},
% and it can make this distinction since it knows that
% \racket{(lambda (x) x)} can not be a function that takes 2 parameters.
% However, the code on Figure~\ref{code:racket:or/c:non-working} would fail
% before running, since the \racket{or/c} contract is unable to know
% which of the two function contracts to use.
% Ideally, this last example should just behave as wrapping
% the function with the \racket{(-> even? even?)} contract.

@@

Racket does provide the \racket{case->} combinator which(partially) supports
overloading. As for the \racket{or/c}, the candidate contracts must have
distinct arities to avoid ambiguity. An example is provided in
Figure~\ref{fig:racket-case-fun}. The resulting possibilities are similar
to static overloading, where one function can take additional parameters for
example (e.g. as supported for Java methods). Unfortunately, it excludes the
overloading of generic operations with fixed arity such as equality, comparison,
arithmetic operators, and so on.

% Racket does provide an alternative for overloading methods, called the
% \racket{case->} contract, however, this overloading of methods only
% works over contracts of functions with different amount of parameters.
% It's a good alternative to allow limited overloading of functions.
% As an example, consider the following non failing piece of code, and
% try to understand how Racket can decide, at each application,
% which function contract to use.

@@

In conclusion, Racket also seems to avoid the difficulties of general unions and
intersections. They make the pragmatic choice of a simple semantics that work
for non higher-order contracts and a limited support for overloading.
% Even if Racket's intent is not to implement union and intersection
% dynamic checks, they do provide contracts that are, in some sense,
% simplified versions of those.
% In our eyes, Racket has accepted the difficulty of correctly and practically
% combine union, intersection, and higher order contracts, and has choose
% to prioritise the highe order ones.

@@

% LocalWords:  Felleisen JSON TOML YAML templating Jsonnet Dhall MyPy
% LocalWords:  Kubernetes DevOps natively nixpkgs disjunctions
% LocalWords:  TypeScript serializable serialization coinductively
% LocalWords:  coinduction coinductive

>>>>

Racket is a language based on the Scheme dialect of Lisp. Among established
languages, Racket is probably the one with the most comprehensive contract
system~\cite{RacketContracts}. Regarding union and intersections, Racket provides
the \racket{and/c} and \racket{or/c} combinators for contract.

The
intersection combinator \racket{and/c} corresponds to the naive
interpretation of intersection described in
Section~\ref{sec:flat-and-inter}: applying contract \racket{(and/c A
  B)} is like applying contract \racket{A} then applying contract \racket{B}. In
particular \racket{and/c} doesn't model overloading. As in
Figure~\ref{fig:naive-impl}, the example given in
Figure~\ref{fig:racket-overloading} always fail because no argument satisfies
both \racket{number?} and \racket{string?}.

@@

The union combinator \racket{or/c} is similarly simple. It must be
able to decide immediately which branch holds: \racket{or/c} is a
simple Boolean disjunction. For instance, when higher-order contracts are combined using \racket{or/c}, Racket imposes
that contracts must be distinguishable by their arity. Doing so, there is at most
one candidate that can be selected directly. This is illustrated in
Figure~\ref{code:racket:or/c:working} whose program is accepted. On the other
hand, the program of Figure~\ref{code:racket:or/c:non-working} is rejected.

@@

To compensate for the fact that \racket{and/c} doesn't support
overloading,
Racket provides a second intersection-like combinator: \racket{case->}. As for the \racket{or/c}, the candidate contracts must have
distinct arities to avoid ambiguity, unlike \racket{or/c} the
alternative chosen when the function is called rather than when the
contract is applied to the function. An example is provided in
Figure~\ref{fig:racket-case-fun}. The resulting possibilities are similar
to static overloading, where one function can take additional parameters for
example (e.g. as supported for Java methods). On the other hand, it excludes the
overloading of generic operations with fixed arity such as equality, comparison,
arithmetic operators, and so on.

@@

In conclusion, Racket, a programming language with a large user base,
avoids the difficulties of general unions and intersections (Section~\ref{sec:issues-sem}). They make
the pragmatic choice of a simple semantics with limited support for
higher-order contracts in intersections and unions.

@@

% LocalWords:  Felleisen JSON TOML YAML templating Jsonnet Dhall MyPy
% LocalWords:  Kubernetes DevOps natively nixpkgs disjunctions
% LocalWords:  TypeScript serializable serialization coinductively
% LocalWords:  coinduction coinductive combinator combinators arities
% LocalWords:  disjunction
